Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline

### 2017-09-13 draft/response_to_referee.txt

As mentioned earlier, such definition of S/N has been adopted within the HSC SSP collaboration, and has
been used for multiple technical and scientific purposes.  Therefore, we think the S/N plotted in Fig 4 and 
Fig 9 are the most meaningful and useful ones for HSC data user.

We slightly modify the text to explain the definition of S/N and flux error better: 
    - In Section 5.1.1, we explain the PSF flux error in more details.
    - In Section 5.2.1, we explain that for extended objects modeled by CModel, definition of S/N is not as
      straightforward as point sources.  Here we choose to use the CModel flux and flux error to simply
      demonstrate the typical "significancy" of galaxies at fixed magnitude, and show that hscPipe can
      relibaly detect and measure flux for galaxies down to certain magnitude.  We also mention that
      uncertainty of CModel flux is dominated by systematics, and here the flux error only summarises the
      statistical ones.  Therefore, our tests should provide more realistic estimation of average flux error
      (including both systematic and statistical ones) for galaxy at certain magnitude.

-----
# 6. Pure statistical errors in figures.

We agree with the referee that it is very useful to compare with the statistical uncertainties. 
However, as we explained for comment #5, "pure statistical random errors" based on photon 
statistics are not very practical to get from hscPipe.

Instead, the flux error from hscPipe for both PSF and CModel photometry are relied on the
detailed variance information that summarize most of the statistical errors. These flux or
magnitude error are also provided to the HSC data user as the most frequently used
information on photometric uncertainties, it is useful to compare them with the results of
our SynPipe tests.

On Fig 5, 6, 7, 10, 11, 12, and 15, we have overplot a pair of grey dashed lines that
outline the running median of statistical uncertainties measured by the hscPipe (in
positive and negative form).  For magnitude, we simply use the magnitude error provided by
hscPipe, for color, we add the magnitude uncertainties for both bands in quadrature.

Given the realistic situation, we think these comparisons cover the requirements by the
referee, although are not exactly the same.  We did not specifically mentioned these
comparesions because the results are exactly as we expected based on the algorithms
themselves: for PSF photometry, the statistical uncertainties from hscPipe are similar but
still slightly lower compared to the median uncertainties from SynPipe test; For CModel,
the statistical flux errors from hscPipe are smaller compared to the uncertainties from
SynPipe test as systematic uncertainties (model-dependent errors) dominate the real error
budget of CModel photometry.

We now briefly mention these comparisons in each section to draw readers attention to
these information.

@@

-----
# 7. Precision and accuracy of PSF magnitude

We believe this comment is also covered by the previous one.  For PSF magnitude, the
statistical uncertainties obtained by hscPipe is similar but still slightly lower than the
photometric uncertainties based on the SynPipe test.  As mentioned, due to the correlated
noise from the image warping and coadding process, the statistical uncertainties of PSF
flux measured by hscPipe is likely to be over-optimistic.

Relevant text has been added to Section 5.1.2 and 5.1.3.

@@

-----
# 8. p. 7-8: "High level reduction pipeline": as stated more clearly in the hscPipe technical paper (Bosch et
al. 2017). Low-level reduction involves basic image detrending and calibration; high-level processes involves
the ones that generate coadded images and science-ready catalogs.  Now use the same explanation here to be
consistent with the technical paper and provide more details.

-----
# 9. "purity and completeness"

We agree with the referee that the purity and completeness tests are very important for
survey like this, and SynPipe can be a useful tool for this purpose.

At the same time, careful tests for this purpose are beyond the scope of this paper, where
we simply want to demonstrate the application of SynPipe and provide simple photometric
benchmark tests for stars and galaxies.  Following the suggestions from the referee, we
add more discussions about this in the discussion section.  And here are the main points:
    - Detection limits and completeness are normally defined for point sources.  As we
      show using the PSF S/N, it depends on the seeing condition and show spatial changes 
      across the 
    - Detection completeness of point sources for HSC DR1 is investigated in Aihara+2017
      (see Section 5.7) using a different method. 
    - The efficiency 
    -

@@

----
1. p.7: should be 'mas' not 'max', corrected.

----
2. p.10, about 'rectangular cutout': the images are stored in a 2-D array, hence has a rectangular nature to
it.  And, it is much easier to convolve the model and shift the 2-D rectangular array before we add them to
the HSC coadd images.  When we generate the model image array using GalSim, we try to make sure that all fluxes 
of the model is contained in the image. For galaxies, we cutout the model at 10 x Re of the Sersic model. This 
is fine for most galaxies, but still miss small fraction of light for galaxy with very high Sersic index.
First, we try to avoide model with very high Sersic index in the input catalog, as they are often problematic
models for real galaxies; Second, given the magnitude distribution of the model galaxies (most have i>23.0
mag), the results we are showing here are not sensitive to the flux distributions outside 10 x Re.  
The referee is correct that estimation of R_e could be sensitive to the very outskirt when Sersic index is
high, but we'd like to point out that the CModel method used in hscPipe does not use any pixel with
S/N < 5, and only provides n=4 model fits (dev) to the "footprint" of the galaxy.  We do not expect accurate
R_e estimates for high Sersic index galaxies in the beginning.

We add a footnote to briefly explain this choice.

----
3. p.11, about "noise":

It is Poisson noise, and we state that clearly in the text.

----
4. p.13:

We think the location and context of this figure makes it very clear that it is for the input
parameters.  Now we state that clearly in the caption.

----
5. Fig. 4 and 5: has been corrected.

----
6. p.22:

For stars, we think the likely reasons for this are: 
    - There are actually very small fraction of stars with those extreme colors, which

----
7.

----
8. P.29

@@

Yes, it should be (i-z) color, and has been corrected.

----
9. Astrometric Calibration Fig 14

HSC image has pixel scale of 0.168 arcsec/pixel.  We use arcsec as unit since when evaluating the astrometric
calibration for HSC DR1, arcsec or mas are used in Aihara+2017.  Information about pixel scale is added to the
caption of the figure and the text.

----
10. the definition of highly blended

The referee is correct, that we should use consistent definition for highly blended
objects, which is b>0.05.  We have make it consistent across the draft.

---- 
11. Table 1, 2, 3, and 4

### 2017-09-13 draft/ms_v5.tex
Most major comments

    The precision of PSF magnitude for the \texttt{badSeeing} \tract{} shows similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the precision slightly degrades. 
    The \texttt{badSeeing} \tract{} has ${\sim}4$\% precision at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% precision at
    $i_{\mathrm{Input}}{\sim}25.0$ mag.
    \citet{HSCDR1} evaluates the precision of PSF magnitude via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds it at 1-2\% level,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the precision of 
    the HSC photometry down to our detection limit\footnote{Strictly speaking, the 
    precision reported here should be considered as upper limits because 
    \synpipe{} currently does not consider the systematic uncertainties in PSF 
    modeling. But this effect should be subdominant.}

@@

	The statistical uncertainties of \forced{} PSF colors are summarized in 
	Table \ref{tab:psfcolor}

@@

    For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm that is an improved 
    version of the SDSS \cmodel{} photometry.
    For details about the \cmodel{} algorithm, please see description in Section 
    \ref{ssec:hscpipe} and Bosch\etal (in prep.).
    Despite the limitations of the \cmodel{} method (e.g., sensitivity to the 
    background subtraction and to deblending failures), it can deliver robust 
    PSF-corrected fluxes and colors for galaxies.

@@

	The precisions and accuracies of \forced{} \cmodel{} magnitudes shown here are 
	summarized in Table \ref{tab:cmodelmag}

@@

	The statistical uncertainties of \forced{} \cmodel{} colors are summarized in 
	Table \ref{tab:cmodelcolor}

@@

    Star--galaxy separation becomes increasingly difficult at faint magnitudes 
    ($i<24.0$ mag).
    In \hscpipe{}, the star--galaxy classification primarily depends on the 
    extendedness value, which is measured by the magnitude difference between PSF and 
    \cmodel{} photometry in $i$-band.
    For more details about this algorithm, please see Bosch\etal (in prep.).

@@

        \item The performance of the current star--galaxy separation algorithm in 
            \hscpipe{} is not perfect, and degrades at fainter magnitudes and with 
            poor seeing conditions.
            The fraction of stars that are misclassified as extended objects is still
            quite high. 
            $>20$\% of the synthetic stars at $i> 22.5$ mag are misclassified
            \footnote{Note that this does not mean high fraction of contamination 
            in the galaxy sample. Please see Bosch\etal (in prep.) for more 
            details.}

@@

    To evaluate the degree to which an object is blended with others, \hscpipe{} 
    introduces the blendedness parameter: $b$ (\texttt{[grizy]blendedness\_abs\_flux}).
    We briefly define  $b$ below; please refer to Bosch\etal (in prep.) and Murata\etal 
    (in prep.) for more details.

@@

    The current version of \hscpipe{} contains a bug related to the calculation of $b$
    parameter. 
    For bright object (e.g. $i<23.0$ mag), its effect can be ignored.  
    For fainter object, it mostly leads to underestimate of $b$ parameter by $< 12$\%, 
    especially for objects with $-3.5 < \log_{10} b < -1.5$. 
    This bug has been fixed for the future release of \hscpipe{} and HSC survey data. 
    For the \synpipe{} tests here, it does not qualitatively change the conclusions 
    related to the highly-blended objects and about the relationship between $b$ and
    photometric precision. 
    Please see Murata\etal (in prep.) and Bosch\etal (in prep.) for more details.

>>>>

    The precision of PSF magnitude for the \texttt{badSeeing} \tract{} shows similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the precision slightly degrades. 
    The \texttt{badSeeing} \tract{} has ${\sim}4$\% precision at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% precision at
    $i_{\mathrm{Input}}{\sim}25.0$ mag.
    \citet{HSCDR1} evaluates the precision of PSF magnitude via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds it at 1-2\% level,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the precision of 
    the HSC photometry down to our detection limit\footnote{Strictly speaking, the 
    precision reported here should be considered as upper limits because 
    \synpipe{} currently does not consider the systematic uncertainties in PSF 
    modeling. But this effect should be subdominant.}
    \song{
    On Fig~\ref{fig:psf_mag}, we also compare the \synpipe{} test results with the 
    median \psf{} magnitude error obtained by \hscpipe{} in each bin after taking 
    the aperture correction (see Section~4.9.2 in \citealt{Bosch2017}) uncertainty 
    into account (pairs of grey long-dashed lines on each plot). 
    We find that, for \psf{} photometry, the flux errors from \hscpipe{} are very 
    close to what we estimate using \synpipe{} once the aperture correction 
    error is considered.
    Without the aperture correction error, it seems that \hscpipe{} underestimate 
    the statistical uncertainties of \psf{} magnitude, most likely due to the 
    correlated noises introduced during the image warping and stacking step.
    }

@@

	The statistical uncertainties of \forced{} PSF colors are summarized in 
	Table \ref{tab:psfcolor}. 
	\song{
	On Fig~\ref{fig:psf_color}, we also compare the \synpipe{} results with the 
	uncertainties of \psf{} color obtained by \hscpipe{} (pair of grey long-dashed 
	lines). 
	Here, we simply add the magnitude errors from both bands in quadrature. 
	We find that the average uncertainties of \psf{} color from \hscpipe{} are 
	consistent with the results of \synpipe{} test. 
	Although we should emphasize again that the uncertainties of PSF models are not 
	taken into account. 
    }

@@

    For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm that is an improved 
    version of the SDSS \cmodel{} photometry.
    For details about the \cmodel{} algorithm, please see description in Section 
    \ref{ssec:hscpipe} and \citet{Bosch2017}.
    Despite the limitations of the \cmodel{} method (e.g., sensitivity to the 
    background subtraction and to deblending failures), it can deliver robust 
    PSF-corrected fluxes and colors for galaxies.

@@

	The precisions and accuracies of \forced{} \cmodel{} magnitudes shown here are 
	summarized in Table \ref{tab:cmodelmag}. 
	\song{
	Similar to \psf{} photometry, here we also compare the average \cmodel{} magnitude 
	error provided by \hscpipe{} in each magnitude bin with our \synpipe{} test results
	(pairs of grey long-dashed lines on each plot). 
	The \cmodel{} magnitude error in \hscpipe{} only considers the statistical 
	uncertainties based on the variance information. 
	As expected, the real error budget for \cmodel{} magnitude is dominated by 
	systematics in the model fitting process, and the \cmodel{} flux uncertainties 
	from \hscpipe{} are underestimated. 
	We stress that user of \cmodel{} photometry should take this into consideration.
	}

@@

	The statistical uncertainties of \forced{} \cmodel{} colors are summarized in 
	Table \ref{tab:cmodelcolor}. 
	\song{
	We also compare the \cmodel{} color uncertainties estimated by \hscpipe{} with 
	our \synpipe{} results. 
	We still simply add the magnitude error from each band in quadrature.
	For $(g-i)$ color, we see that the statistical uncertainties of \cmodel{} colors 
	are very similar to the \synpipe{} results. 
	This is consistent with the design of the \forced{} \cmodel{} algorithm, where 
	precisions are expected to be much better for colors than for fluxes themselves 
	(see Section 4.9.9 of \citealt{Bosch2017}).
	For $(i-y)$ color, we can draw similar conclusion except for the faint end 
	($i>24.0$ mag), where the uncertainties from \synpipe{} test become higher than 
	the statistical uncertainties from \hscpipe{}. 
	This is likely due to the scattered light in $y$--band that has not been accounted 
	for (see Section~5.8.14 in \citealt{HSCDR1}). 
	}

@@

    Star--galaxy separation becomes increasingly difficult at faint magnitudes 
    ($i<24.0$ mag).
    In \hscpipe{}, the star--galaxy classification primarily depends on the 
    extendedness value, which is measured by the magnitude difference between PSF and 
    \cmodel{} photometry in $i$-band.
    For more details about this algorithm, please see \citet{Bosch2017}.

@@

        \item The performance of the current star--galaxy separation algorithm in 
            \hscpipe{} is not perfect, and degrades at fainter magnitudes and with 
            poor seeing conditions.
            The fraction of stars that are misclassified as extended objects is still
            quite high. 
            $>20$\% of the synthetic stars at $i> 22.5$ mag are misclassified
            \footnote{Note that this does not mean high fraction of contamination 
            in the galaxy sample. Please see \citet{Bosch2017} for more 
            details.}

@@

    To evaluate the degree to which an object is blended with others, \hscpipe{} 
    introduces the blendedness parameter: $b$ (\texttt{[grizy]blendedness\_abs\_flux}).
    We briefly define  $b$ below; please refer to \citet{Bosch2017} and Murata\etal 
    (in prep.) for more details.

@@

    The current version of \hscpipe{} contains a bug related to the calculation of $b$
    parameter. 
    For bright object (e.g. $i<23.0$ mag), its effect can be ignored.  
    For fainter object, it mostly leads to underestimate of $b$ parameter by $< 12$\%, 
    especially for objects with $-3.5 < \log_{10} b < -1.5$. 
    This bug has been fixed for the future release of \hscpipe{} and HSC survey data. 
    For the \synpipe{} tests here, it does not qualitatively change the conclusions 
    related to the highly-blended objects and about the relationship between $b$ and
    photometric precision. 
    Please see Murata\etal (in prep.) and \citet{Bosch2017} for more details.
